A Mountaintop View Requires Minimal Sorting: \break An Instance Optimal Contour Tree Algorithm

### 2014-10-27 paint-contour-tree.tex
more changes to section 6, 7

\newcommand{\build}{{\tt build}}
\newcommand{\col}{col}
\newcommand{\fcol}{full}
\newcommand{\cur}{cur}
\newcommand{\cut}{{\tt cut}}
\newcommand{\cont}{\psi}
\newcommand{\cost}{\mathop{cost}}
\newcommand{\h}{att}
\newcommand{\jc}{\cJ_C}
\newcommand{\lift}{{\tt lift}}
\newcommand{\mcol}{mcol}
\newcommand{\merge}{{\tt merge}}
\newcommand{\onestep}{{\tt piece}}
\newcommand{\paint}{{\tt paint}}
\newcommand{\pal}{P}
\newcommand{\pmax}{P_{\max}}
\newcommand{\rain}{{\tt rain}}
\newcommand{\reeb}{\cC}
\newcommand{\redH}{\widetilde{H}}
\newcommand{\rep}{rep}
\newcommand{\st}{K}
\newcommand{\surgery}{{\tt surgery}}
\newcommand{\touch}{T}
\newcommand{\update}{{\tt update}}
\newcommand{\wet}{{\tt wet}}

@@

The main theorem of this section asserts that the contours trees of minimum dominant manifolds have a simple description.
This description is based on \emph{join} and \emph{split} trees, as given by~\cite{csa-cctad-00}.
Conventially, all edges are directed from higher to lower function value.
We first explain what these trees are.

\begin{definition} \label{join} The join tree $\cJ(\MM)$ of $\MM$ is built on vertex set $V(\MM)$.
The directed edge $(u,v)$ is present when $u$ is the smallest valued vertex in a connected component of $\MM^+_v$
\emph{and} $v$ is connected to this component (in $\MM$). The split tree $\cS(\MM)$ is obtained
by looking at $\MM^-_v$ (or alternately, by taking the join tree of the inversion of $\MM$).

The critical join tree $\jc(\MM)$ is built on $V(\MM) \setminus M$. 
The directed edge $(u,v)$ is present when $u$ is the smallest valued vertex in $V(\MM) \setminus M$ in a connected component of $\MM^+_v$
and $v$ is connected to this component (in $\MM$). 
\end{definition}

Some basic facts about these trees. 
%Note that unlike the contour tree, all vertices in $\MM$ are present in $\cJ(\MM)$ and $\cS(\MM)$. 
All outdegrees in $\cJ(\MM)$ and all indegrees in $\cS(\MM)$ are at most $1$.
Indegree $2$ vertices in $\cJ(\MM)$ are joins, and outdegree $2$ vertices in $\cS(\MM)$ are splits.
Leaves in $\cJ(\MM)$ are maxima, and leaves in $\cS(\MM)$ are minima.
The key observation is that $\cS(\MM)$ is trivial for minimum dominant $\MM$.

\begin{lemma} \label{lem:split} $\cS(\MM)$ consists of:
\begin{asparaitem}
	\item A single path (in sorted order) with all vertices except non-dominant minima.
	\item Each non-dominant minimum is attached to a unique split.
\end{asparaitem}
\end{lemma}

@@

@@

@@

The main algorithmic contribution is a new algorithm for computing join trees of any triangulated simplicial
complex $\MM$.

As part of the initialization, we determine
all critical points in $\MM$. For any vertex $v$, we can look at $f(v) \pm \delta$-contours 
in $f|_{B_\eps(v)}$ to determine the up and down degrees. This determines the criticality (and type)
of $v$. This requires $O(|\MM|)$ time, since every facet is looked at proportional to the number of its
vertices.

{\bf Painting:} The central tool is a notion 
of \emph{painting} of $\MM$. Initially associate a color with each maximum. Imagine there being a large
can of paint of a distinct color at each maxima $x$. We will spill different paint from each maximum and watch it flow down.
This is analogous to the raining in the previous section, but paint is a much more viscous liquid.
\emph{So paint only flows down edges, and it does not color the interior of facets.} Furthermore, paints
do not mix, so every edge of $\MM$ gets a unique color. This process (and indeed the entire algorithm)
works purely on the 1-skeleton of $\MM$, which is just a graph.
%
\begin{definition} \label{def:paint} Let the 1-skeleton of $\MM$ have edge set $E$ and maxima $X$.
A  \emph{painting} of $\MM$ is a map $\chi:X \cup E \mapsto [|X|]$ with the following property. 
 Consider an edge $e$. There exists a descending path from some maximum $x$ to $e$
	consisting of edges in $E$, such that all edges along this path have the same color as $x$. (The maximum $x$
	is well-defined to be the unique maximum with the same color as $e$.)

An \emph{initial} painting has the additional restriction that the restriction $\chi:X \mapsto [|X|]$ is a bijection.
\end{definition}
%
Paintings are not unique. It is straightforward to construct an initial painting of $\MM$ in $O(|\MM|)$ time, 
using a descending BFS from each maximum that does not explore previously colored edges. 
As the algorithm proceeds, the bijection property will not hold, though we always maintain a valid painting.

{\bf Colors of vertices:} Note that a painting only colors edges, not vertices. We associate certain
sets of colors with vertices.

\begin{definition} \label{def:color-set} Fix a painting $\chi$, and a color $c$.
A vertex $v$ is \emph{touched by $c$} if $x$ is incident to a $c$-colored
edge with $v$ at the lower endpoint. For $v$, $\col(v)$ is the set of colors that touch $v$.
\end{definition}

Abusing notation, we say a contour $\phi$ is of color $c$ if all edges that intersect $\phi$ have color $c$.

\begin{definition} \label{def:color-full} Consider vertex $v$. An \emph{up-star} of $v$ is the set 
of edges that all connected to a fixed component of $\MM^+_v$. (Each $v$
participates in at mot $2$ up-stars. A color $c \in \col(v)$ \emph{fully touches} $v$
if all edges in an up-star are colored $v$.
%and let $\phi$ be a $(f(v)+\delta)$-contour in
%$f|_{B_\eps(v)}$. If $\phi$ is monochromatic of color $c$, then $v$ is \emph{fully touched by $c$}. The set of colors
%that fully touch $v$ is denoted $\fcol(v)$.
\end{definition}

%Note that $\fcol(v)$ has size at most $2$, by the Morse condition. 
Note that $\col(v)$ 
%and $\fcol(v)$ can
be initially determined in $O(|\MM|)$ time, by a simple linear scan.

@@

We first list out the basic data structures used, and then summarize our approach.

@@

We initialize as follows. For each critical vertex $v$ and from each of its up-stars, we pick an arbitrary color $c$ touching $v$. 
We insert $v$ into $T(c)$. Observe that $v$ participates in at least $1$ and at most $2$ heaps.

\medskip
\noindent
{\bf The union-find data structure on colors:} We will repeatedly perform unions
of classes of colors, and this will be maintained as a standard union-find data structure.
For any color $c$, $\rep(c)$ denotes the representative of its class. We initialize
by simply setting $\rep(c) = c$ for all colors.

@@

\medskip
\noindent
{\bf Attachment vertex $\h(c)$:} For each color $c$, we maintain a critical point $\h(c)$ of this color.
We will maintain the guarantee that the portion of the contour tree above (and including) $\h(c)$ has already been constructed.
This is initialized to the maximum corresponding to $c$.

\medskip
As the algorithm proceeds, it is convenient to imagine
it maintaining a set of contours $\Phi = \{\phi_1, \phi_2, \ldots, \phi_k\}$, which are contained
in incomparable edges of $\cC(\MM)$. Suppose we cut $\MM$ by each of these contours, and took
the simplicial complex $\NN$ where $\phi^-_i$'s are all boundary maxima. The algorithm will
have already determined the contour trees of the simplicial complices
where $\phi^+_i$ is a boundary maxima.

@@

While this is the intuitive viewpoint, we describe how this information is actually maintained.
The vertices $\{\h(\rep(c))\}$ represent the set of contours $\Phi$. In terms of the contour tree, 
we have already constructed the contour trees rooted at each $h(\rep(c))$. This is trivially true at the initialization,
when each $h(c)$ is a maximum. One can now treat each $h(\rep(c))$ as a maximum, corresponding to (a possibly new) color $\rep(c)$.
To reduce confusion between the original painting and the new colors, we will assume that $\col(v)$ and $\fcol(v)$
are as given by the original painting.

@@

@@

\smallskip
\begin{compactenum}
	\item While there are unprocessed critical points:
	\begin{compactenum}
		\item Run $\update(\st)$. Pop $\st$ to get $h$.
		\item Let $\cur(h) = \{\rep(c) | c \in \col(h)\}$.
		\item For all $c' \in \cur(h)$:
		\begin{compactenum}
			\item Add edge $(\h(c'),h)$ to $\cC(\MM)$.
			\item Delete $h$ from $T(c')$ (if present).
		\end{compactenum}
	\item Merge heaps $\{T(c') | c' \in \cur(h)\}$.
	\item Take union of $\cur(h)$ and denote resulting color by $\widehat{c}$.
	\item Set $\h(\widehat{c}) = h$ and mark $h$ as processed.
	\end{compactenum}
\end{compactenum}
\end{minipage}}

@@

A few simple facts:
\begin{asparaitem}
	\item When $\update(\st)$ terminates, the head $h$ is ripe. 
	\item After $v$ is processed, all edges incident to $v$ has the same color.
\end{asparaitem}

@@

@@

\begin{claim} \label{clm:ripe} For ripe vertex $v$, $\{\h(\rep(c)) | c \in \col(v)\}$ is exactly
the set of parents of $v$.
\end{claim}

\begin{proof} \Sesh{proof is a little sketchy.}
Suppose $v$ is ripe. For any $c$ touching $v$, $v$ is the highest in $T(\rep(c))$.
Consider some such $\rep(c)$, for an edge connecting to component $\PP$ of $\MM^+_v$.
Either all edges in $\PP$ are colored $\rep(c)$ or there exists critical vertex $w \in \PP$
fully touched by $\rep(c)$ and another color. In the latter case, $w$ has not been processed,
so $w \in T(\rep(c))$ (contradiction). Therefore, all edges in $\PP$ are colored $\rep(c)$.
Suppose there is some critical vertex in $\PP$. If it is not processed yet, it will
be in $T(\rep(c))$, which is not possible. Thus, all critical vertices in $\PP$ have been processed.

>>>>

\newcommand{\build}{{\tt build}}
\newcommand{\col}{col}
\newcommand{\fcol}{full}
\newcommand{\cur}{cur}
\newcommand{\cut}{{\tt cut}}
\newcommand{\cont}{\psi}
\newcommand{\cost}{\mathop{cost}}
\newcommand{\h}{att}
\newcommand{\init}{{\tt init}}
\newcommand{\jc}{\cJ_C}
\newcommand{\lift}{{\tt lift}}
\newcommand{\mcol}{mcol}
\newcommand{\merge}{{\tt merge}}
\newcommand{\onestep}{{\tt piece}}
\newcommand{\paint}{{\tt paint}}
\newcommand{\pal}{P}
\newcommand{\pmax}{P_{\max}}
\newcommand{\rain}{{\tt rain}}
\newcommand{\reeb}{\cC}
\newcommand{\redH}{\widetilde{H}}
\newcommand{\rep}{rep}
\newcommand{\st}{K}
\newcommand{\surgery}{{\tt surgery}}
\newcommand{\touch}{T}
\newcommand{\update}{{\tt update}}
\newcommand{\wet}{{\tt wet}}

@@

The main theorem of this section asserts that the contours trees of minimum dominant manifolds have a simple description.
The exact statement will require some definitions and notation.
We required the notions of \emph{join} and \emph{split} trees, as given by~\cite{csa-cctad-00}.
Conventially, all edges are directed from higher to lower function value. We will use the terminology
``parent" for an adjacent node that is higher. Because of the orientation, a node
may have multiple parents.
%We first explain what these trees are.

\begin{definition} \label{join} The join tree $\cJ(\MM)$ of $\MM$ is built on vertex set $V(\MM)$.
The directed edge $(u,v)$ is present when $u$ is the smallest valued vertex in a connected component of $\MM^+_v$
\emph{and} $v$ is connected to this component (in $\MM$). The split tree $\cS(\MM)$ is obtained
by looking at $\MM^-_v$ (or alternately, by taking the join tree of the inversion of $\MM$). 
\end{definition}

Some basic facts about these trees. 
%Note that unlike the contour tree, all vertices in $\MM$ are present in $\cJ(\MM)$ and $\cS(\MM)$. 
All outdegrees in $\cJ(\MM)$ are at most $1$, all indegree $2$ vertices are joins, all leaves are maxima,
and the global minimum is the root. All indegrees in $\cS(\MM)$ are at most $1$, all outdegree $2$
vertices are splits, leaves are minima, and the global maximum is the root.
The key observation is that $\cS(\MM)$ is trivial for minimum dominant $\MM$.

\begin{lemma} \label{lem:split} $\cS(\MM)$ consists of:
\begin{asparaitem}
	\item A single path (in sorted order) with all vertices except non-dominant minima.
	\item Each non-dominant minimum is attached to a unique split (which is adjacent to it).
\end{asparaitem}
\end{lemma}

@@

It is convenient to denote the non-dominant minima as $m_1, m_2, \ldots, m_k$
and the corresponding splits (as given by the lemma above) as $s_1, s_2, \ldots, s_k$.

@@

\begin{definition} The critical join tree $\jc(\MM)$ is built on the set $V'$ of all
critical points that are \emph{not} non-dominant minima. 
The directed edge $(u,v)$ is present when $u$ is the smallest valued vertex in $V'$ in a connected component of $\MM^+_v$
and $v$ is connected to this component (in $\MM$). 
\end{definition}

@@

The main algorithmic contribution is a new algorithm for computing join trees of any triangulated simplicial
complex $\MM$.
%
%
%
%As part of the initialization, we determine
%all critical points in $\MM$. For any vertex $v$, we can look at $f(v) \pm \delta$-contours 
%in $f|_{B_\eps(v)}$ to determine the up and down degrees. This determines the criticality (and type)
%of $v$. This requires $O(|\MM|)$ time, since every facet is looked at proportional to the number of its
%vertices.

{\bf Painting:} The central tool is a notion 
of \emph{painting} $\MM$. Initially associate a color with each maximum. Imagine there being a large
can of paint of a distinct color at each maxima $x$. We will spill different paint from each maximum and watch it flow down.
This is analogous to the raining in the previous section, but paint is a much more viscous liquid.
\emph{So paint only flows down edges, and it does not color the interior of facets.} Furthermore, paints
do not mix, so every edge of $\MM$ gets a unique color. This process (and indeed the entire algorithm)
works purely on the 1-skeleton of $\MM$, which is just a graph.
%
\begin{definition} \label{def:paint} Let the 1-skeleton of $\MM$ have edge set $E$ and maxima $X$.
A  \emph{painting} of $\MM$ is a map $\chi:X \cup E \mapsto [|X|]$ with the following property. 
 Consider an edge $e$. There exists a descending path from some maximum $x$ to $e$
	consisting of edges in $E$, such that all edges along this path have the same color as $x$.

An \emph{initial} painting has the additional property that the restriction $\chi:X \mapsto [|X|]$ is a bijection.
\end{definition}
%
%Paintings are not unique. It is straightforward to construct an initial painting of $\MM$ in $O(|\MM|)$ time, 
%using a descending BFS from each maximum that does not explore previously colored edges. 
%As the algorithm proceeds, the bijection property will not hold, though we always maintain a valid painting.

A painting only colors edges, not vertices. We associate certain
sets of colors with vertices.

\begin{definition} \label{def:color-set} Fix a painting $\chi$ and vertex $v$.
\begin{asparaitem}
	 \item An \emph{up-star} of $v$ is the set of edges that all connected to a fixed component of $\MM^+_v$.
	 \item A vertex $v$ is \emph{touched by color $c$} if $x$ is incident to a $c$-colored
	 edge with $v$ at the lower endpoint. For $v$, $\col(v)$ is the set of colors that touch $v$.
	 \item A color $c \in \col(v)$ \emph{fully touches} $v$ if all edges in an up-star are colored $v$.
\end{asparaitem}
\end{definition}

Each non-maxima $v$ participates in at least $1$ and at most $2$ up-stars (the latter iff
$v$ is a join).

\ignore{
Abusing notation, we say a contour $\phi$ is of color $c$ if all edges that intersect $\phi$ have color $c$.

\begin{definition} \label{def:color-full} Consider vertex $v$. An \emph{up-star} of $v$ is the set 
of edges that all connected to a fixed component of $\MM^+_v$. 
%and let $\phi$ be a $(f(v)+\delta)$-contour in
%$f|_{B_\eps(v)}$. If $\phi$ is monochromatic of color $c$, then $v$ is \emph{fully touched by $c$}. The set of colors
%that fully touch $v$ is denoted $\fcol(v)$.
\end{definition}

%Note that $\fcol(v)$ has size at most $2$, by the Morse condition. 
Note that $\col(v)$ 
%and $\fcol(v)$ can
be initially determined in $O(|\MM|)$ time, by a simple linear scan.
}

@@

We give an informal description of the algorithm, which might motivate the data structures.

Our algorithm incrementally builds $\jc(\MM)$ from the leaves (maxima) to the root (dominant minimum). 
Let us focus on an initial painting, where each colors has 1-1 correspondence with the maxima.
Consider two sibling leaves $\ell_1, \ell_2$ and their common parent $v$. The leaves are maxima,
and $v$ is a join that ``merges" $\ell_1, \ell_2$. A nice mental picture is given when $\MM$ is 2-dimensional.
In that case, there are ``mounds" corresponding to each $\ell_1$ and $\ell_2$ that merge
at a valley $v$. Suppose this was the entire input, and $\ell_1$ was colored blue and $\ell_2$ was colored red. 
Both the mounds are colored completely blue or red, while $v$ is touched by both colors.
Moreover, each up-star of $v$ (one connecting to each mound) is fully colored. So this
indicates that $v$ joins the blue maxima and red maxima in $\jc(\MM)$.

This is precisely how we hope to exploit the information in the painting. We prove later that
when some join $v$ has both up-stars fully colored, the corresponding maxima (of those colors) are exactly
the parents of $v$ in $\jc(\MM)$. To proceed further, we ``merge" the colors red and blue into a new color, purple. In other words,
we replace all red and blue edges by purple edges. This indicates that the red and blue maxima have been
handled. Imagine flattening the red and blue mounds till $v$, so now that erstwhile join $v$
is a new maxima. Suppose purple paint was poured from $v$. In terms on $\jc(\MM)$, this
is equivalent to remove leaves $\ell_1$ and $\ell_2$, now making $v$ as a leaf.
Alternately, $\jc(\MM)$ has been constructed up to $v$, and it remains to determine $v$'s parent.

Of course, things are more complicated when there are other mounds. There may be a yellow mound,
corresponding to $\ell_3$ that joins with the blue mound higher up (at vertex $u$). In $\jc(\MM)$, $\ell_2$ and $\ell_3$
are sibling leaves, and $\ell_1$ is a sibling of some ancestor of these leaves. So we cannot
merge red and blue, until yellow and blue merge. Naturally, we use priority queues to handle this issue.
We know that $u$ must also be (fully) touched by blue. So all critical vertices touched by blue
are put into a priority queue keyed by height, and vertices are handled in that order.

What happens when finally blue and red join at $v$? We merge the two colors, but now have
a blue heap and red queues of critical vertices. We also need to merge the priority queues
to get a consistent painting. And that decides the use of \emph{binomial heaps} to store the
critical points of one color. Merges can be done in logarithmic time, the same time
required for deletes.

In this discussion, we have ignored a somewhat annoying problem. Up-stars of critical vertices
may actually be touched by numerous colors, not just one as we assumed above. A simple solution
would be to insert vertices into heaps corresponding to all colors touching it. But there could
be super-constant numbers of copies of a vertex, and handling all these copies would lead
to an extra overhead. We show that it suffices to simply put each vertex $v$ into at most two heaps,
one for each up-star. We are guaranteed that when $v$ needs to processed, its up-stars
automatically have one color, because of all the color merges that previously occured.

With this preamble, we are ready to discuss our data structures. We discuss all the initializations
in the next section.

@@

%We initialize as follows. For each critical vertex $v$ and from each of its up-stars, we pick an arbitrary color $c$ touching $v$. 
%We insert $v$ into $T(c)$. Observe that $v$ participates in at least $1$ and at most $2$ heaps.
%

\medskip
\noindent
{\bf The union-find data structure on colors:} We will repeatedly perform unions
of classes of colors, and this will be maintained as a standard union-find data structure.
For any color $c$, $\rep(c)$ denotes the representative of its class.

@@

\medskip
\noindent
{\bf Attachment vertex $\h(c)$:} For each color $c$, we maintain a critical point $\h(c)$ of this color.
We will maintain the guarantee that the portion of the contour tree above (and including) $\h(c)$ has already been constructed.
%This is initialized to the maximum corresponding to $c$.

\ignore{
\medskip
As the algorithm proceeds, it is convenient to imagine
it maintaining a set of contours $\Phi = \{\phi_1, \phi_2, \ldots, \phi_k\}$, which are contained
in incomparable edges of $\cC(\MM)$. Suppose we cut $\MM$ by each of these contours, and took
the simplicial complex $\NN$ where $\phi^-_i$'s are all boundary maxima. The algorithm will
have already determined the contour trees of the simplicial complices
where $\phi^+_i$ is a boundary maxima.

@@

While this is the intuitive viewpoint, we describe how this information is actually maintained.
The vertices $\{\h(\rep(c))\}$ represent the set of contours $\Phi$. In terms of the contour tree, 
we have already constructed the contour trees rooted at each $h(\rep(c))$. This is trivially true at the initialization,
when each $h(c)$ is a maximum. One can now treat each $h(\rep(c))$ as a maximum, corresponding to (a possibly new) color $\rep(c)$.
To reduce confusion between the original painting and the new colors, we will assume that $\col(v)$ and $\fcol(v)$
are as given by the original painting. 
}

@@

\medskip
\fbox{
\begin{minipage}{0.9\textwidth}
{\bf $\init(\MM)$}

\smallskip
\begin{compactenum}
	\item Construct an initial painting of $\MM$ using a descending BFS from maxima that does not explore previously colored edges.
	\item Determine all critical points in $\MM$. For each $v$, look at $(f(v) \pm \delta)$-contours in $f|_{B_\eps(v)}$ 
	to determine the up and down degrees. 
	\item Mark each critical $v$ as unprocessed.
	\item For each critical $v$ and each up-star, pick arbitrary color $c$ touching $v$. Insert $v$ into $T(c)$.
	\item Initialize $\rep(c) = c$ and set $\h(c)$ to be the unique maximum colored $c$.
\end{compactenum}
\end{minipage}}

@@

\smallskip
\begin{compactenum}
	\item Run $\init(\MM)$.
	\item While there are unprocessed critical points:
	\begin{compactenum}
		\item Run $\update(\st)$. Pop $\st$ to get $h$.
		\item Let $\cur(h) = \{\rep(c) | c \in \col(h)\}$.
		\item For all $c' \in \cur(h)$:
		\begin{compactenum}
			\item Add edge $(\h(c'),h)$ to $\jc(\MM)$.
			\item Delete $h$ from $T(c')$ (if present).
		\end{compactenum}
	\item Merge heaps $\{T(c') | c' \in \cur(h)\}$.
	\item Take union of $\cur(h)$ and denote resulting color by $\widehat{c}$.
	\item Set $\h(\widehat{c}) = h$ and mark $h$ as processed.
	\end{compactenum}
\end{compactenum}
\end{minipage}}

@@

\bigskip

A few simple facts:
\begin{asparaitem}
	\item At all times, the colors form a valid painting.
	\item Each vertex is present in at most $2$ heaps. After processing, it is removed from all heaps.
	\item After $v$ is processed, all edges incident to $v$ have the same color (technically, same representative). 
	\item When $\update(\st)$ terminates, the head $h$ is ripe. 
\end{asparaitem}

@@

Our main workhorse is the following technical lemma.

We prove a series of claims that will lead to the correctness proof.

\begin{claim} \label{clm:upstar} Consider ripe vertex $v$ and take the up-star connecting
to some component of $\MM^+_v$. All edges in this component and the up-star have the same color.
\end{claim}

\begin{proof} Let $c$ be the color of some edge in this up-star
By ripeness, $v$ is the highest in $T(\rep(c))$.
Denote the component of $\MM^+_v$ by $\PP$.
By \Lem{full}, either all edges in $\PP$ are colored $\rep(c)$ or there exists critical vertex $w \in \PP$
fully touched by $\rep(c)$ and another color. In the latter case, $w$ has not been processed,
so $w \in T(\rep(c))$ (contradiction to ripeness). Therefore, all edges in $\PP$ are colored $\rep(c)$.
\end{proof}

\begin{claim} \label{clm:process} The partial output on the processed vertices is exactly
the restriction of $\jc(\MM)$ to these vertices.
\end{claim}

\begin{proof} More generally, we prove the following: all outputs on processed vertices
are edges of $\cC(\MM)$ and for any color $c$, $\h(c)$ is the lowest processed vertex
of that color. We prove by induction on the processing order. Initially, this is trivially
true, handling these base case. For the induction step, consider the situation
when $v$ is being processed.

The vertex $v$ is ripe. Take any up-star, and the corresponding component $\PP$
of $\MM^+_v$ that it connects to. By \Clm{upstar}, all edges in $\PP$ and the up-star
have the same color (say $c$). If some critical vertex in $\PP$ is not processed,
it must be in $T(c)$, which violates the ripeness of $v$.
Thus, all critical vertices in $\PP$ have been processed.

@@
